Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-02 neurips-20/cvx-flats.tex
Meeting notes

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff r \in \gamma(p)$, since $r \in \gamma(p)$ implies $\ell(r;p) = \risk{\ell}(p) := \inf_{r' \in \R} \exploss{\ell}{r'}{p}$, and if $r \not \in \gamma(p)$, then $\ell(r;p) > \risk{\ell}(p)$.
  \raft{We don't define $\risk{\ell}$}\jessiet{Added in line, but very subtle.  not a huge fan.}
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Then
  \raft{Discuss proof}
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, we have that $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:cvx-flats}
%	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
%	For all $r\in\gamma(p)$, if there is no flat $F \subseteq \affhull(\simplex)$ \jessie{affine hull or affine span?} of codimension at most $d+1$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
%	
	If there is a convex surrogate loss $L : \reals^d \times\Y \to \reals$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
\end{theorem}
\begin{proof}
%	We prove the contrapositive: if there is a convex surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d+1$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
%	
	Let $\Gamma := \prop{L}$ and suppose $(L, \psi)$ indirectly elicits $\gamma$ for some link $\psi : \reals^d \to \R$, so that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix $p \in \simplex$, and take any $u \in \Gamma(p)$ such that $\psi(u) = r \in \gamma(p)$.
	We will show $F \cap \simplex \subseteq \Gamma_u$; the result then follows from $\Gamma_u \subseteq \gamma_r$.
	As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p} \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.
  % (Observe that in order for $L$ to be convex, $L(\cdot, y)$ must be convex for all $y \in \Y$.)

	Expanding, we have $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
	In order for $\vec 0$ to be in this $p$-weighted Minkowski sum, each choice of $x_i$ is a flat for $\partial L(u, i)$.
  \raft{What is a flat for a convex function?}
	Defining the matrix $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$, we have $W p = \sum_y p_y x_y = \vec 0$
  (cf. $\mathbf{A}^m$ from~\cite[Theorem 16]{ramaswamy2016convex}).
	Now taking $F := \ker(W) \cap \affhull(\simplex)$, we have $\codim(F) = \rank(W) \leq d$, and $p \in F$ by construction.
  To see $F \cap \simplex \subseteq \gamma_r$, from the chain of equivalences above, we have for any $q\in\simplex$ that $q \in \ker W \implies u \in \Gamma(q) \implies q \in \Gamma_u \subseteq \gamma_r$.

  \raf{Old:}
	That would mean that $q \in F \implies Wq = \vec 0 \implies \sum_y p_y x_y^i = 0$ for all $i \in [d]$, which in turn implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u \subseteq \gamma_{r}$.
\end{proof}

@@

\begin{lemma}\label{lem:affhull-relint}
	Let $C\subseteq\reals^m$ be convex.
	Then for any affine subspace $F\subseteq\affhull(C)$ with $F\cap\relint C \neq \emptyset$, we have $\affhull(F\cap C) = F$.
\end{lemma}
\begin{proof}
	As $\affhull(F) = F$ and $F\cap C\subseteq F$, the inclusion $\affhull(F\cap C) \subseteq F$ is clear.
	For the reverse, let $p\in F\cap\relint C$ and let $B\subseteq C$ be a relatively open set containing $p$.
	For any $q\in F \subseteq \affhull(C)$, we thus have $q' = p + \epsilon (q-p) \in B$ for sufficiently small $\epsilon > 0$.
	\bo{How does this follow exactly from B relatively open?}
	As $q' = (1-\epsilon) p + \epsilon q$, we have $q' \in \affhull(F)\cap C = F\cap C$.
	As $q = (1-1/\epsilon) p + (1/\epsilon) q'$, we thus have $q\in\affhull(F\cap C)$.
\end{proof}

>>>>

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff r \in \gamma(p)$, since $r \in \gamma(p)$ implies $\ell(r;p) = \risk{\ell}(p) := \inf_{r' \in \R} \exploss{\ell}{r'}{p}$, and if $r \not \in \gamma(p)$, then $\ell(r;p) > \risk{\ell}(p)$.
  \raft{We don't define $\risk{\ell}$}\jessiet{Added in line, but very subtle.  not a huge fan.}
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Then \btw{todo: emphasize that it's target loss}
%  \raft{Discuss proof}
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, we have that $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:cvx-flats}
%	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
%	For all $r\in\gamma(p)$, if there is no flat $F \subseteq \affhull(\simplex)$ \jessie{affine hull or affine span?} of codimension at most $d+1$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
%	
	If there is a convex surrogate loss $L : \reals^d \times\Y \to \reals$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
\end{theorem}
\begin{proof}
%	We prove the contrapositive: if there is a convex surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is a flat $F \subseteq \affhull(\simplex)$ of codimension at most $d+1$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.
%	
	Let $\Gamma := \prop{L}$ and suppose $(L, \psi)$ indirectly elicits $\gamma$ for some link $\psi : \reals^d \to \R$, so that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix $p \in \simplex$ and $r \in \gamma(p)$, and take any $u \in \Gamma(p)$ such that $\psi(u) = r \in \gamma(p)$.\jessie{Does such a $u$ exist?}
	We will show $F \cap \simplex \subseteq \Gamma_u$; the result then follows from $\Gamma_u \subseteq \gamma_r$.
	As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p} \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.
  % (Observe that in order for $L$ to be convex, $L(\cdot, y)$ must be convex for all $y \in \Y$.)

	Expanding, we have $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
%	In particular, $u \in \Gamma(p)$ if and only if $\vec 0$ is a $p$-convex combination of subgradients.
	Defining the matrix $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$, we have $W p = \sum_y p_y x_y = \vec 0$
  (cf. $\mathbf{A}^m$ from~\cite[Theorem 16]{ramaswamy2016convex}).
	Now taking $F := \ker(W) \cap \affhull(\simplex)$, we have $\codim(F) = \rank(W) \leq d$, and $p \in F$ by construction.
  To see $F \cap \simplex \subseteq \gamma_r$, from the chain of equivalences above, we have for any $q\in\simplex$ that $q \in \ker W \implies u \in \Gamma(q) \implies q \in \Gamma_u \subseteq \gamma_r$.

  \raf{Old:}
	That would mean that $q \in F \implies Wq = \vec 0 \implies \sum_y q_y (x_y)_j = 0$ for all $j \in [d]$, which in turn implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u \subseteq \gamma_{r}$.
\end{proof}

@@

\begin{lemma}\label{lem:affhull-relint}
	Let $C\subseteq\reals^m$ be convex.
	Then for any affine subspace $F\subseteq\affhull(C)$ with $F\cap\relint C \neq \emptyset$, we have $\affhull(F\cap C) = F$.
\end{lemma}
\begin{proof}
	As $\affhull(F) = F$ and $F\cap C\subseteq F$, the inclusion $\affhull(F\cap C) \subseteq F$ is clear.
	For the reverse, let $p\in F\cap\relint C$ and let $B\subseteq C$ be a relatively open set containing $p$.
	For any $q\in F \subseteq \affhull(C)$, we thus have $q' = p + \epsilon (q-p) \in B$ for sufficiently small $\epsilon > 0$.
%	\bo{How does this follow exactly from B relatively open?}
	As $q' = (1-\epsilon) p + \epsilon q$, we have $q' \in \affhull(F)\cap C = F\cap C$.
	As $q = (1-1/\epsilon) p + (1/\epsilon) q'$, we thus have $q\in\affhull(F\cap C)$.
\end{proof}
